Scaling Automated Theory Exploration

### 2015-08-06 AI4FM/article.tex
Tweaks

The \emph{theory exploration} (TE) paradigm provides software support
for traditional Mathematical workflows; namely, deriving ``interesting''
consequences from formal definitions \cite{RISC1482}. Early
implementations like \textsc{Theorema} \cite{buchberger2000theory} emphasised
interactivity, in a similar way to computer algebra systems like
\textsc{Mathematica} (in which \textsc{Theorema} is implemented) or
interactive theorem provers. Subsequent systems have investigated
\emph{automated} theory exploration, for tasks such as lemma discovery
\cite{Hipster}.

@@

We are implementing a system called \textsc{ML4HS} to investigate scalable
automated theory exploration for Haskell. The current form of \textsc{ML4HS} is
a pre-processor for \textsc{QuickSpec} which attempts to make exploration of
large theories more tractable by splitting them into a collection of smaller
theories: the divide and conquer approach mentioned above. This is
similar to reducing the search space in automated theorem proving via premise
selection \cite{kuhlwein2012overview}. An overview of \textsc{ML4HS} is given in
\autoref{ml4hsdia}. We have integrated our pre-processor with the \textsc{Cabal} and
\textsc{Nix} package managers so that \textsc{ML4HS} can use regular Haskell
packages as theories.

>>>>

The \emph{theory exploration} (TE) paradigm provides software support
for traditional Mathematical workflows; namely, deriving ``interesting''
consequences from formal definitions \cite{RISC1482}. Early
implementations like \textsc{Theorema} \cite{buchberger2000theory} provided
interactive environments, similar to computer algebra systems like
\textsc{Mathematica} (in which \textsc{Theorema} is implemented) or
interactive theorem provers. Subsequent systems have investigated
\emph{automated} theory exploration, for tasks such as lemma discovery
\cite{Hipster}.

@@

We are implementing a system called \textsc{ML4HS} to investigate scalable
automated theory exploration for Haskell. The current form of \textsc{ML4HS} is
a pre-processor for \textsc{QuickSpec} which attempts to make exploration of
large theories more tractable by splitting them into a collection of smaller
theories: the divide and conquer approach mentioned above. This is
similar to reducing the search space in automated theorem proving via premise
selection \cite{kuhlwein2012overview}. The basic idea of \textsc{ML4HS} is shown
in \autoref{ml4hsdia}: theories, in the form of Haskell packages, are split up
and sent to concurrent instances of QuickSpec, and the results collated.
